---
title: "Lab7"
author: "Leo Lu"
date: "2024-10-17"
output: "github_document"
---
# Load packages and dataset
```{r}
library(haven)
library(bruceR)
library(dplyr)
library(ggstatsplot)
library(performance)
library(sjPlot)

ex <- read_sav("/Users/leolu/Documents/GitHub/Lab7/ex.sav")

lab7<-read.csv("/Users/leolu/Documents/GitHub/Lab7/lab7.csv")



```

# Correlation
```{r}
#First, you will need to select variables or composites of interest

ex <- ex %>%
  select(Conscientiousness, AverageHoursofSleep, GPA)

#The Corr function also gives you a correlation plot between all variables, but it doesn't look pretty enough and not very customizable

Corr(ex)

#If you hate the correlation plot design, you can use the following function
ggcorrmat(ex)

#If you want to visualize the correlation between two specifcic variables, you can use the following code
ggplot(ex, aes(x = Conscientiousness, y = GPA)) + geom_point() + geom_smooth() + theme_bruce()

#Note that the line will naturally follow the data points, but you can specify to plot the fittest line by adding method = lm to the geom_smooth() function
ggplot(ex, aes(x = Conscientiousness, y = GPA)) + geom_point() + geom_smooth(method = lm) + theme_bruce()




```

# Regression
```{r}
#step 1: build a model
model<-lm(GPA ~ Conscientiousness + AverageHoursofSleep, data = ex)
#GPA preicted by conscientiousness + averageHoursofSleep

#step 2: check the assumptions
check_model(model)

#step 3: summarize results; add ,std = TRUE if you want to get standardized coefficients
model_summary(model)


# or using the tab_model functionl; add  ,show.std = TRUE if you want to get standardized coefficients
tab_model(model)

#step 4: plot the results; change type ="est" to type = "std" if you want to plot the standardized coefficients
plot_model(model,  type ="est",  show.values = TRUE, vline.color = "#1B191999", line.size = 1.5, dot.size = 2.5, colors = "blue") + theme_bruce()

```



# Q1: Use the Corr or the ggcorrmat function to plot the correlation between all variables. Which personality traits are correlated with satisfaction with life and which one is not? 
```{r}

Corr(lab7)

ggcorrmat(lab7)

ggplot(lab7, aes(x = Extraversion, Agreeableness, Conscientiousness, EmotionalStability, Openness, y = SWL)) + geom_point() + geom_smooth() + theme_bruce()

ggplot(lab7, aes(x = Extraversion, Agreeableness, Conscientiousness, EmotionalStability, Openness, y = SWL)) + geom_point() + geom_smooth(method = lm) + theme_bruce()

#All variables except openness are correlated with satisfaction wth life.

```

# Q2: Run a multiple regression using the Big 5 personality traits to predict satisfaction with life and plot the results
```{r}
#for this lab assignment you would assume assumptions are met, but for your own analysis, you will need to examine assumptions carefully

model<-lm(SWL ~ Extraversion + Agreeableness + Conscientiousness + EmotionalStability + Openness, data = lab7)

check_model(model)

model_summary(model)

tab_model(model)

plot_model(model,  type ="est",  show.values = TRUE, vline.color = "#1B191999", line.size = 1.5, dot.size = 2.5, colors = "blue") + theme_bruce()


```

# Q3: Interpret the intercept, the meaning of each of the unstandardized coefficient (e.g., one unit increase in A is associated with x point increase in B), and the R2. Which personality trait(s) can explain unique variance in satisfaction with life when controlling for each other, and which ones cannot? If someone asks you for advice on how to improve life satisfaction, based on your results, which personality trait would you recommend them to change and why? 

The intercept is 2.191. When all traits are 0, the SWL is 2.191 

One unit increase in Extraversion is associated with a 0.14 increase in Satisfaction With Life. 
One unit increase in Agreeableness is associated with a 0.062 increase in Satisfaction With life. 
One unit increase in Conscientiousness is associated with a 0.019 increase in Satisfaction With Life. 
One unit increase in Emotional Stability is associated with a 0.134 increase in Satisfaction With Life.
One unit increase in Openness is associated with a -0.101 decrease in Satisfaction With Life.

As for r^2, 13.8% of the variability is explained by the whole model.

Extraversion, Emotional Stability, and Openness are able to explain unique variances while Agreeableness and conscientiousness are not.

If someone asks me which personality traits I would advise on improving life satisfaction. I would highly recommend Extraversion as it causes the largest increase in Life satisfaction. Then I will recommend Emotional Stability as it is the only other trait that positively predicts Life satisfaction. I would not mention Agreeableness and Conscientiousness as they have no significant impact on life satisfaction. I would definitely not recommend Openness as positively increasing Openness lowers Life satisfaction.